Individual and Differential Harm in Redistricting

### 2022-05-25 paper/paper.Rmd
correlation appendix

The harm framework combines two key ideas, which are motivated at greater length in Section \ref{sec:motivation}.
First, we operate at the level of the individual voter, and examine whether each voter is represented by the candidate of their choice.
Doing so both unifies our approach to gerrymandering across contexts and also reflects a core goal of representative democracy: that elected officials represent the preferences of the voting public [cites on why districts are good].

@@

New Jersey is just one state, and while some of the correlations between measures here are found in other states, some are not.
Appendix \ref{app:corr} calculates these correlations across a number of states.
The most consistent correlation is that between our estimate of differential partisan harm and the expected number of seats won by each party, which is always above \_\_ in the states we study.

@@

```{=tex}
\vspace{1in}
\appendix
\renewcommand\thefigure{\thesection.\arabic{figure}}
\setcounter{figure}{0}
```
# Computing Harm Measures from Precinct-level Election Data {#app:formulas}

@@

The table below reports the correlation between differential harm and the partisan gerrymandering metrics used in Section \ref{sec:nj} for \_\_ states which cover a wide range of partisan leans and number of seats.
The simulations are provided by the 50-State Simulation Project [@mccartan2022fiftystates].

>>>>

The harm framework combines two key ideas, which are motivated at greater length in Section \ref{sec:motivation}.
First, we operate at the level of the individual voter, and examine whether each voter is represented by the candidate of their choice.
Doing so both unifies our approach to gerrymandering across contexts and also reflects a core goal of representative democracy: that elected officials represent the preferences of the voting public to the greatest extent possible [@brunell2010; other cites on why districts are good. Federalist No. 52?].

@@

New Jersey is just one state, and while some of the correlations between measures here are found in other states, some are not.
Appendix \ref{app:corr} calculates these correlations across eight states.
The most consistent correlation is that between our estimate of differential partisan harm and the expected number of seats won by each party, which is usually around 0.98, and always above 0.66 in the eight states we study.

@@

```{=tex}
\vspace{1in}
\appendix
\renewcommand\thefigure{\thesection.\arabic{figure}}
\renewcommand\thetable{\thesection.\arabic{table}}
\setcounter{figure}{0}
\setcounter{table}{0}
```
# Computing Harm Measures from Precinct-level Election Data {#app:formulas}

@@

Table \ref{tab:corr} below reports the correlation between differential harm and the partisan gerrymandering metrics used in Section \ref{sec:nj} for eight states which cover a wide range of partisan leans and number of seats:

- Utah, with 4 districts and a Democratic vote share baseline of 33%
- Mississippi, with 4 districts and a Democratic vote share baseline of 43%
- Iowa, with 4 districts and a Democratic vote share baseline of 46%
- North Carolina, with 14 districts and a Democratic vote share baseline of 50%
- Wisconsin, with 8 districts and a Democratic vote share baseline of 51%
- New Jersey, with 12 districts and a Democratic vote share baseline of 57%
- Massachusetts, with 9 districts and a Democratic vote share baseline of 61%
- California, with 52 districts and a Democratic vote share baseline of 65%

The correlations are calculated across 5,000 simulated plans for each state.
The simulations are provided by the 50-State Simulation Project [@mccartan2022fiftystates].
In particular, the New Jersey simulations are not the same as those used in Section \ref{sec:nj}.

\begin{table}[h!]
\centering
\input{figures/corr_table.tex}
\vspace{0.1in}
\caption{Correlations between five gerrymandering metrics and partisan differential harm, for simulated districting plans from eight states. Declination is NA for Massachusetts since there are no Republican-leaning districts at baseline under any simulated plan.}
\label{tab:corr}
\end{table}
